---
title: "Final codedown"
author: "Magnus"
date: "9/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Loading the relevant libraries

```{r Load Libraries, include = FALSE}
pacman::p_load('tidyverse', 'ggplot2', 'lme4', 'lmerTest', 'MuMIn', 'pastecs', 'purr')

```

### Loading data

```{r Load Data, include = FALSE}
#we wrote the code together line by line meaning we all contributed to all the code equally

#load data
df <- read_csv('autisme.csv')

#remove meaningless column
df$X1 <- NULL

#remove rows where CHI_MLU = N/A
df <- df %>% filter(!is.na(CHI_MLU))

```


### Characterize the participants (Exercise 1)

Identify relevant variables: participants demographic characteristics, diagnosis, ADOS, Verbal IQ, Non Verbal IQ, Socialization, Visit, Number of words used, Number of unique words used, mean length of utterance in both child and parents.

Make sure the variables are in the right format.

Describe the characteristics of the two groups of participants and whether the two groups are well matched.

```{r descriptive stats, include = TRUE}

#see dataframe and the variables 
str(df)

#changing the variables and getting an overview of the characteristics 
df %>% 
    mutate(
    Gender = as.factor(Gender),
    SUBJ = as.factor(SUBJ),
    Ethnicity = as.factor(df$Ethnicity),
    Diagnosis = as.factor(Diagnosis)
  ) %>% 
  split(df$Diagnosis) %>%
  map(summary)

df <- df %>% 
    mutate(
    Gender = as.factor(Gender),
    SUBJ = as.factor(SUBJ),
    Ethnicity = as.factor(df$Ethnicity),
    Diagnosis = as.factor(Diagnosis)
  )

```

## Let's test hypothesis 1: Children with ASD display a language impairment  (Exercise 2)

### Hypothesis: The child's MLU changes: i) over time, ii) according to diagnosis

Let's start with a simple mixed effects linear model

Remember to plot the data first and then to run a statistical test.
- Which variable(s) should be included as fixed factors? Visit and Diagnosis
- Which variable(s) should be included as random factors? Subject ID

```{r ex2, include = FALSE}

#plotting the data

ggplot(df) + 
  aes(x = Age, y = CHI_MLU, color = Diagnosis) +
  geom_point() + geom_smooth(method = 'lm', se = F) +
  theme_light()

ggplot(df) + 
  aes(x = VISIT, y = CHI_MLU, color = Diagnosis) +
  geom_point() + geom_smooth(method = 'lm', se = F) +
  theme_light()

df %>% 
  filter(VISIT == 1 | VISIT == 6) %>% 
  ggplot(aes(x = Diagnosis, y = CHI_MLU, fill = Diagnosis)) +
  geom_boxplot(stat = "boxplot", color = "grey", outlier.colour = "black", outlier.shape = 16, outlier.size = 2, notch = FALSE) +
  stat_summary(fun.y = mean, geom = "point", shape = 23, size = 4) + 
  theme_linedraw() + 
  facet_wrap(~ VISIT)

```


```{r}
#creating models

null <- lmerTest::lmer(CHI_MLU ~ 1 + (1|SUBJ), df, REML = F)
m1 <- lmerTest::lmer(CHI_MLU ~ VISIT + (1|SUBJ), df, REML = F)
m2 <- lmerTest::lmer(CHI_MLU ~ Diagnosis + (1|SUBJ), df, REML = F)
m3 <- lmerTest::lmer(CHI_MLU ~ VISIT + Diagnosis + (1|SUBJ), df, REML = F)
m4 <- lmerTest::lmer(CHI_MLU ~ VISIT + Diagnosis + (1 + VISIT|SUBJ), df, REML = F)
m5 <- lmerTest::lmer(CHI_MLU ~ VISIT * Diagnosis + (1 + VISIT|SUBJ), df, REML = F)

```

Regarding the initially stated hypotheses, children with ASD do display a language impairment. This can be seen from the interaction effect, 


How would you evaluate whether the model is a good model?

```{r ex2 evaluate, include = FALSE}

#anova
anova(null, m1, m2, m3, m4, m5)

#mumin
MuMIn::r.squaredGLMM(m1)
MuMIn::r.squaredGLMM(m2)
MuMIn::r.squaredGLMM(m3)
MuMIn::r.squaredGLMM(m4)
MuMIn::r.squaredGLMM(m5)

```


Not too good, right? Let's check whether a growth curve model is better.
Remember: a growth curve model assesses whether changes in time can be described by linear, or quadratic, or cubic (or... etc.) components.

First build the different models, then compare them to see which one is better.

```{r ex2 growth curve, include = FALSE}

#adding quadratic component
m5_sq <- update(m5, .~. + I(VISIT^2))

#failed to converge, making new model
m6 <- lmerTest::lmer(CHI_MLU ~ VISIT * Diagnosis + (1 | SUBJ), df, REML = F)

#updating m6
m6_sq <- update(m6, .~. + I(VISIT^2))

#is it better than m6 without quadratic component?
anova(m6, m6_sq)

#is it better than our best model from exercise 2?
anova(m6_sq, m5)

```

Exciting right? Let's check whether the model is doing an alright job at fitting the data. Plot the actual CHI_MLU data against the predictions of the model fitted(model). 

```{r}
#getting the residuals
sum_m6_sq <- summary(m6_sq)
residuals <- sum_m6_sq$residuals

#making a df with actual values and predicted values
exciting <- data.frame("actual" = df$CHI_MLU, "residuals" = sum_m6_sq$residuals, "predicted" = df$CHI_MLU + sum_m6_sq$residuals)

#plotting actual vs predicted
ggplot(exciting) + aes(x = actual, y = predicted) + geom_point() + geom_smooth(method = "lm")

```

## Let's test hypothesis 2: Parents speak equally to children with ASD and TD  (Exercise 3)

### Hypothesis: Parental MLU changes: i) over time, ii) according to diagnosis

```{r ex3, include = FALSE}

mot1 <- lmerTest::lmer(MOT_MLU ~ VISIT + Diagnosis + (1|SUBJ), df, REML = F)


summary(mot1)

```


### Adding new variables (Exercise 4)

Your task now is to figure out how to best describe the children linguistic trajectory. The dataset contains a bunch of additional demographic, cognitive and clinical variables (e.g.verbal and non-verbal IQ). Try them out and identify the statistical models that best describes your data (that is, the children's MLU). Describe how you selected the best model and send the code to run the model to Victor and Byurakn.


```{r ex4, include = FALSE}
#checking for correlation between different variables
cor.test(df$types_MOT, df$tokens_MOT, method = 'spearman')
cor.test(df$MOT_MLU, df$tokens_MOT, method = 'spearman')
cor.test(df$MOT_MLU, df$types_MOT, method = 'spearman')
cor.test(df$MullenRaw, df$ExpressiveLangRaw, method = 'spearman')
cor.test(df$ExpressiveLangRaw, df$types_CHI, method = 'spearman')
cor.test(df$MullenRaw, df$types_CHI, method = 'spearman')
cor.test(df$Age, df$VISIT, method = 'spearman')
cor.test(df$types_MOT, df$VISIT, method = 'spearman')

#testing if socializaiton is a significant variable
df_new <- df %>% filter(!is.na(Socialization))
m7_new <- lmerTest::lmer(CHI_MLU ~ VISIT + Diagnosis + types_CHI + MOT_MLU + Socialization + (1 + VISIT|SUBJ), df_new, REML = F)
m8_new <- lmerTest::lmer(CHI_MLU ~ VISIT + Diagnosis + types_CHI + MOT_MLU + (1 + VISIT|SUBJ), df_new, REML = F)
anova(m7, m8)
#judging by the output socialization is not a valuable predictor

#finding the best model
m7 <- lmerTest::lmer(CHI_MLU ~ VISIT + Diagnosis + types_CHI + MOT_MLU + (1 + VISIT|SUBJ), df, REML = F)

m8 <- lmerTest::lmer(CHI_MLU ~ VISIT + Diagnosis * types_CHI + MOT_MLU + (1 + VISIT|SUBJ), df, REML = F)

m9 <- lmerTest::lmer(CHI_MLU ~ VISIT + Diagnosis + types_CHI + (1 + VISIT|SUBJ), df, REML = F)

m10 <- lmerTest::lmer(CHI_MLU ~ VISIT + Diagnosis + (1 + VISIT|SUBJ), df, REML = F)

anova(m7, m8, m9, m10)

summary(m7)

```